---
title: "Fast Univariate Inference under Survey Design"
format: 
  html:
    code-fold: true
execute:
  warning: false
  echo: true
bibliography: references.bib
---

------------------------------------------------------------------------

## What We Have Done

1.  **FUI Analysis on NHATS Data**:

    -   Explored the association between diurnal physical activity and covariates using Fast univariate inference proposed by Erjia [@cui2022].

2.  **NHATS introduction and motivation**

    -   **Processed NHATS Round 11 and 12 accelerometry data:**
        -   Extracted weights, PSUs, and strata from survey design variables.
        -   Designed survey objects:
            -   Single Observation: Incorporated strata, PSU, and weights to reduce bias.
            -   Multiple Observations: Used weights only due to limitations in multi-stage sampling probabilities.
    -   **NHATS (Round 11 and 12)** provides accelerometry data for a nationally representative sample of Medicare beneficiaries aged 71 and older.

3.  **Implementation of Weighted Functional Univariate Inference**:

    -   **Framework Development:** Extended the Fast Univariate Inference (FUI) framework proposed by Erjia et al. to accommodate complex survey designs.

    -   **Two scenarios:**

        -   **Single Observation**: No random effects, analyzed using `svyglm` [@svylme]

        -   **Multiple Observations**: Includes random effects, analyzed using `svy2lme` [@lumley2023]

            -   **Notes:** For `svy2lme`, separate probabilities or weights for each stage of sampling are required. However, NHATS does not provide such detailed information. Therefore, we only utilized available weights for the analysis.

            -   As a comparison, **WeMix** (Weighted Mixed Effects Model) was explored as it similarly handles weighted linear mixed effects models.

4.  **Three-Step Inferential Procedures**:

    1.  **Point-wise Model Fitting**:

        -   Fit separate models at each location $s_l \in S$:
            -   Single observation: Used `svyglm` to include strata, PSU, and weights
            -   Multiple observations: Used `svy2lme` to include weights (no strata/PSU due to data limitations)
        -   **Extended bootstrap methods:**
            -   Current methods include:
                -   Rao-Wu-Yue-Beaumont Bootstrap
                -   Non-parametric Bootstrap
                -   Preston Bootstrap
                -   Finite Population Bootstrap (FPB)
            -   Planned addition: Unified Bootstrap for stratified cluster sampling [@wang2022]
                -   The code is not publicly available. But the original code has been obtained from the author. You can contact me if needed.

    2.  **Smoothing Along the Functional Domain**:

        -   Applied penalized splines to obtain smooth estimates of functional coefficients:
            -   Used `mgcv::gam` to smooth fixed effect estimates $\beta(s)$ along the domain.

    3.  **Joint Confidence Bands (JCB)**:

        -   Applied fPCA to analyze variance structure of Bootstrap samples.
        -   Used Supremum Norm to derive critical values and construct JCBs.

5.  **Simulation summary**:

    -   **Single Point Simulations:** Verify the combination between LMER/svy2lme and RWYB bootstap

        -   [Simple Random Sampling (SRS)]{.underline}: Data generated without stratification or clustering; weights are uniform across all observations

        -   [Stratified Multi-Stage Sampling]{.underline}: Incorporates strata, PSUs per stratum, and varying weights to simulate real-world survey designs

        -   [Complex Design-Dependent Covariates]{.underline}: Covariates and weights depend explicitly on strata and PSU structures, modeling real-world sampling complexities.

        -   **Preliminary Results**
            -   RWYB Bootstrap achieves the most consistent coverage across all scenarios.
            -   Non-Parametric Bootstrap is faster but underperforms in complex designs.

    -   **\[Current Work\] Functional Domain Simulations**

        -   Single-level functional data
        -   Multi-level functional data

# Analysis Examples (for NHATS)

```{r,warning=FALSE, results = 'hide', collapse=TRUE}
library(here)
library(devtools)
library(fastFMM)
library(readxl)
library(readr)
library(haven)
library(dplyr)
library(survey)
library(svrep)
library(progress)
library(svylme)
library(lme4)
library(mgcv)
library(WeMix)
library(ggplot2)
library(gridExtra)

source('wFUI.R')
source('functions.R')
source('plot_fui_ci.R')
```

#### NHATS data loading (100 subjects)

```{r,warning=FALSE, results = 'hide', collapse=TRUE}
## NHATS minute-level Actigraph data
# dat.raw = read_dta(here('data/NHATS_Round_11_Accel_Det_File.dta'))
# colnames(dat.raw) = sub('ag11dmean', '', colnames(dat.raw))
# dat = dat.raw |>
#   rename(subjectid = spid)
# colnames(dat)[5:1444] = sub("min", 'Y.', colnames(dat)[5:1444])
# 
# ## Sample Design
# dat.cov = read.csv(here('data/NHATS_Round_11_Accel_30Min_Covariates_Weights.csv')) |>
#   dplyr::select(-age, -gender, -race) |>
#   dplyr::rename(subjectid = spid,
#          race = rl5dracehisp,
#          gender = r5dgender,
#          age = r11d2intvrage) |>
#   dplyr::select(subjectid, gender, race, age, 
#          w11anfinwgt0, w11varstrat, w11varunit) |>
#   dplyr::rename(weight = w11anfinwgt0,
#                 psu = w11varunit,
#                 strata = w11varstrat)
#     # weight w#anfinwgt0; 
#     # cluster w#varunit; (=psu)
#     # strata w#varstrat;
# 
# ## Data with Multiple Observations
# dat.m = merge(dat, dat.cov)
# 
# ## Data with Single Observation
# y_columns <- grep("Y.", names(dat), value = TRUE)
# dat.fixed = dat %>%
#   group_by(subjectid) %>%
#   summarise(across(all_of(y_columns), mean, na.rm = TRUE))
# dat.s = merge(dat.fixed, dat.cov)
# 
# ## Save Data
# write_rds(dat.m, here('data', 'NHATS_multiple_obs.rds'))
# write_rds(dat.s, here('data', 'NHATS_single_obs.rds'))
```

```{r,warning=FALSE, results = 'hide', collapse=TRUE}
dat.m = read_rds(here('data', 'NHATS_multiple_obs.rds')) |>
  rename(psu = PSU)
dat.s = read_rds(here('data', 'NHATS_single_obs.rds'))
```

## Single Obs (survey design, no random effects)

### Non-parametric bootstrap

```{r,warning=FALSE, results = 'hide', collapse=TRUE}
# dat.tmp = read_rds(here('data', 'NHATS_single_obs.rds')) |>
#   filter(subjectid %in% dat.s$subjectid[1:100]) |>
#   rename(psu = PSU)
# Y_mat <- dat.tmp[, grepl('Y.', colnames(dat.tmp))]
# dat.fit <- data.frame(Y = Y_mat, 
#                       dat.tmp[,!grepl('Y.', colnames(dat.tmp))])
# formula.FUI = as.formula(paste0('Y~', 'gender'))
# 
# mod.0 <- fui.weight(formula.FUI,
#              analytic = FALSE,
#              argvals = seq(from = 1, to = 1440, by = 30),
#              num_boots = 100,
#              boot_type = 'ref_design',
#              weighted = TRUE,
#              random = FALSE,
#              nknots_min = 10,
#              data = dat.fit)
# 
# export.0 = plot_fui(fuiobj = mod.0,
#          align_x = 1,
#          y_scal_orig = 0.01,
#          x_rescale = 60,
#          return = TRUE)
# 
# write_rds(export.0, here('data', 'mod.0.rds'))
export.0 = read_rds(here('data', 'mod.0.rds'))

fig.0.int = plot_fui_ci(export.0$`(Intercept)`,
            cross_interval = FALSE,
            label = 'Intercept',
            sublabel = 'survey design, no random effects',
            ylim = c(-200, 2000),
            yinterval = 400)

fig.0.gender = plot_fui_ci(export.0$`gender2 FEMALE`,
                           cross_interval = FALSE,
                           label = 'Gender (Female)',
                           sublabel = 'survey design, no random effects',
                           ylim = c(-650, 650),
                           yinterval = 200)

grid.arrange(fig.0.int, fig.0.gender, ncol = 2)
```

### Rao-Wu-Yue-Beaumont

```{r,warning=FALSE, results = 'hide', collapse=TRUE}
# dat.tmp = read_rds(here('data', 'NHATS_single_obs.rds')) |>
#   filter(subjectid %in% dat.s$subjectid[1:100]) |>
#   rename(psu = PSU)
# Y_mat <- dat.tmp[, grepl('Y.', colnames(dat.tmp))]
# dat.fit <- data.frame(Y = Y_mat, 
#                       dat.tmp[,!grepl('Y.', colnames(dat.tmp))])
# formula.FUI = as.formula(paste0('Y~', 'gender'))
# 
# mod.3 <- fui.weight(formula = formula.FUI,
#              analytic = FALSE,
#              argvals = seq(from = 1, to = 1440, by = 30),
#              weighted = TRUE,
#              random = FALSE,
#              boot_type = 'Rao-Wu-Yue-Beaumont',
#              num_boots = 100,
#              nknots_min = 5,
#              data = dat.fit)
# 
# plot_fui(mod.3)
# 
# # Visualization
# export.3 = plot_fui(fuiobj = mod.3,
#          align_x = 1,
#          y_scal_orig = 0.01,
#          x_rescale = 60,
#          return = TRUE)
# 
# write_rds(export.3, here('data', 'mod.3.rds'))
export.3 = read_rds(here('data', 'mod.3.rds'))

fig.3.int = plot_fui_ci(export.3$`(Intercept)`,
            cross_interval = FALSE,
            label = 'Intercept',
            sublabel = 'no survey design, random effects',
            ylim = c(-200, 2000),
            yinterval = 400)

fig.3.gender = plot_fui_ci(export.3$`gender2 FEMALE`,
                           cross_interval = FALSE,
                           label = 'Gender (Female)',
                           sublabel = 'no survey design, random effects',
                           ylim = c(-600, 600),
                           yinterval = 200)

grid.arrange(fig.3.int, fig.3.gender, ncol = 2)
```

## Single Obs (no survey design, no random effects)

### Non-parametric bootstrap

```{r,warning=FALSE, results = 'hide', collapse=TRUE}
# dat.tmp = read_rds(here('data', 'NHATS_single_obs.rds')) |>
#   filter(subjectid %in% dat.s$subjectid[1:100]) |>
#   rename(psu = PSU)
# Y_mat <- dat.tmp[, grepl('Y.', colnames(dat.tmp))]
# dat.fit <- data.frame(Y = Y_mat, 
#                       dat.tmp[,!grepl('Y.', colnames(dat.tmp))])
# formula.FUI = as.formula(paste0('Y~', 'gender'))
# 
# mod.1 <- fui.weight(formula = formula.FUI,
#              analytic = FALSE,
#              argvals = seq(from = 1, to = 1440, by = 30),
#              weighted = FALSE,
#              random = FALSE,
#              boot_type = 'ref_noDesign',
#              num_boots = 100,
#              nknots_min = 5,
#              data = dat.fit)
# 
# export.1 = plot_fui(fuiobj = mod.1,
#          align_x = 1,
#          y_scal_orig = 0.01,
#          x_rescale = 60,
#          return = TRUE)
# 
# write_rds(export.1, here('data', 'mod.1.rds'))
export.1 = read_rds(here('data', 'mod.1.rds'))

fig.1.int = plot_fui_ci(export.1$`(Intercept)`,
            cross_interval = FALSE,
            label = 'Intercept',
            sublabel = 'no survey design, no random effects',
            ylim = c(-200, 2000),
            yinterval = 400)

fig.1.gender = plot_fui_ci(export.1$`gender2 FEMALE`,
                           cross_interval = FALSE,
                           label = 'Gender (Female)',
                           sublabel = 'no survey design, no random effects',
                           ylim = c(-600, 600),
                           yinterval = 200)

grid.arrange(fig.1.int, fig.1.gender, ncol = 2)
```

## Multiple Obs (no survey design, random effects)

### Semiparametric bootstrap

```{r,warning=FALSE, results = 'hide', collapse=TRUE}
# dat.tmp = dat.m |>
#   filter(subjectid %in% dat.s$subjectid[1:100])
# Y_mat <- dat.tmp[, grepl('Y.', colnames(dat.tmp))]
# dat.fit <- data.frame(Y = Y_mat, 
#                       dat.tmp[,!grepl('Y.', colnames(dat.tmp))])
# 
# formula.FUI = as.formula(paste0('Y~gender+', '(1 | subjectid)'))
# 
# mod.2 <- fui.weight(formula.FUI,
#              analytic = FALSE,
#              argvals = seq(from = 1, to = 1440, by = 30),
#              num_boots = 100,
#              num_cores = 10,
#              weighted = FALSE,
#              random = TRUE,
#              boot_type = 'semiparametric',
#              nknots_min = 5,
#              data = dat.fit)
# 


# Visualization
# export.2 = plot_fui(fuiobj = mod.2,
#          align_x = 1,
#          y_scal_orig = 0.01,
#          x_rescale = 60,
#          return = TRUE)
# write_rds(export.2, here('data', 'mod.2.rds'))
export.2 = read_rds(here('data', 'mod.2.rds'))

fig.2.int = plot_fui_ci(export.2$`(Intercept)`,
            cross_interval = FALSE,
            label = 'Intercept',
            sublabel = 'no survey design, random effects',
            ylim = c(-200, 2000),
            yinterval = 400)

fig.2.gender = plot_fui_ci(export.2$`gender2 FEMALE`,
                           cross_interval = FALSE,
                           label = 'Gender (Female)',
                           sublabel = 'no survey design, random effects',
                           ylim = c(-600, 600),
                           yinterval = 200)

grid.arrange(fig.2.int, fig.2.gender, ncol = 2)
```

## Multiple Obs (survey design, random effects)

### Rao-Wu-Yue-Beaumont

```{r,warning=FALSE, results = 'hide', collapse=TRUE}
# dat.tmp = read_rds(here('data', 'NHATS_multiple_obs.rds')) |>
#   filter(subjectid %in% dat.s$subjectid[1:100]) |>
#   rename(psu = PSU)
# Y_mat <- dat.tmp[, grepl('Y.', colnames(dat.tmp))]
# dat.fit <- data.frame(Y = Y_mat, 
#                       dat.tmp[,!grepl('Y.', colnames(dat.tmp))])
# formula.FUI = as.formula(paste0('Y~gender+', '(1 | subjectid)'))
# 
# mod.4 <- fui.weight(formula.FUI,
#              analytic = FALSE,
#              argvals = seq(from = 1, to = 1440, by = 30),
#              num_boots = 100,
#              boot_type = 'Rao-Wu-Yue-Beaumont',
#              weighted = TRUE,
#              random = TRUE,
#              nknots_min = 5,
#              data = dat.fit)
# 
# # Visualization
# export.4 = plot_fui(fuiobj = mod.4,
#          align_x = 1,
#          y_scal_orig = 0.01,
#          x_rescale = 60,
#          return = TRUE)
# 
# write_rds(export.4, here('data', 'mod.4.rds'))
export.4 = read_rds(here('data', 'mod.4.rds'))

fig.4.int = plot_fui_ci(export.4$`(Intercept)`,
            cross_interval = FALSE,
            label = 'Intercept',
            sublabel = 'survey design, random effects',
            ylim = c(-200, 2000),
            yinterval = 400)

fig.4.gender = plot_fui_ci(export.4$`gender2 FEMALE`,
                           cross_interval = FALSE,
                           label = 'Gender (Female)',
                           sublabel = 'survey design, random effects',
                           ylim = c(-600, 650),
                           yinterval = 200)

grid.arrange(fig.4.int, fig.4.gender, ncol = 2)
```

### Preston bootstrap

```{r,warning=FALSE, results = 'hide', collapse=TRUE}
# dat.tmp = read_rds(here('data', 'NHATS_multiple_obs.rds')) |>
#   filter(subjectid %in% dat.s$subjectid[1:100]) |>
#   rename(psu = PSU)
# Y_mat <- dat.tmp[, grepl('Y.', colnames(dat.tmp))]
# dat.fit <- data.frame(Y = Y_mat, 
#                       dat.tmp[,!grepl('Y.', colnames(dat.tmp))])
# formula.FUI = as.formula(paste0('Y~gender+', '(1 | subjectid)'))
# 
# mod.5 <- fui.weight(formula.FUI,
#              analytic = FALSE,
#              argvals = seq(from = 1, to = 1440, by = 30),
#              num_boots = 100,
#              boot_type = 'Preston',
#              weighted = TRUE,
#              random = TRUE,
#              nknots_min = 5,
#              data = dat.fit)
# 
# # Visualization
# export.5 = plot_fui(fuiobj = mod.5,
#          align_x = 1,
#          y_scal_orig = 0.01,
#          x_rescale = 60,
#          return = TRUE)
# 
# write_rds(export.5, here('data', 'mod.5.rds'))
export.5 = read_rds(here('data', 'mod.5.rds'))

fig.5.int = plot_fui_ci(export.5$`(Intercept)`,
            cross_interval = FALSE,
            label = 'Intercept',
            sublabel = 'survey design, random effects',
            ylim = c(-200, 2000),
            yinterval = 400)

fig.5.gender = plot_fui_ci(export.5$`gender2 FEMALE`,
                           cross_interval = FALSE,
                           label = 'Gender (Female)',
                           sublabel = 'survey design, random effects',
                           ylim = c(-600, 650),
                           yinterval = 200)

grid.arrange(fig.5.int, fig.5.gender, ncol = 2)
```

# 03.Results Comparision

```{r,warning=FALSE, results = 'hide', collapse=TRUE}
export.0[[2]]$type = 'Single+Weighted (Incorrectly define design)'
export.1[[2]]$type = 'Single+NonWeighted (Non-parametric)'
export.2[[2]]$type = 'Multiple+Weighted'
export.3[[2]]$type = 'Single+Weighted (RWYB)'
export.4[[2]]$type = 'Multiple+Weighted (RWYB)'
export.5[[2]]$type = 'Multiple+Weighted (Preston)'

combined_export <- bind_rows(export.0[[2]], export.1[[2]], export.2[[2]], 
                             export.3[[2]], export.4[[2]], export.5[[2]])
combined_export$type = as.factor(combined_export$type)

dat = combined_export
colnames(dat) = c('time', 'beta',
                    'lower.pointwise', 'upper.pointwise',
                    'lower.joint', 'upper.joint', 'type')

dat$joint.sig = ifelse(sign(dat$lower.joint) == sign(dat$upper.joint), 1,0)
dat$point.sig = ifelse(sign(dat$lower.pointwise) == sign(dat$upper.pointwise), 1,0)
dat$label.sig = ifelse(dat$joint.sig == 1, 2, dat$point.sig)
```

## Point Esitimate

```{r,warning=FALSE, results = 'hide', collapse=TRUE}
library(RColorBrewer)

cbd.1 = ggplot(dat, aes(x = time / 60, y = beta, color = type, group = type)) +
  geom_hline(yintercept = 0, linetype = "dashed", color = "black") +
  geom_line(size = 1) +  
  scale_x_continuous(
    breaks = c(1, 6, 12, 18, 23),
    labels = function(x) paste0(sprintf("%02d", x), ":00")
  ) +
  scale_color_brewer(palette = "Set2") +
  labs(title = 'Summarized Results',
       x = "Time (hour)", y = 'Physical Intensity (count/min)',
       color = "Type") +
  theme_minimal() +
  theme(
    panel.grid.major = element_line(color = "gray90", size = 0.2),
    axis.text = element_text(size = 12),
    axis.title = element_text(size = 12, face = "bold"),
    legend.title = element_text(size = 12, face = "bold"),
    legend.text = element_text(size = 10)
  )
print(cbd.1)
```

## Point Esitimate + 95% point-wise CI

```{r,warning=FALSE, results = 'hide', collapse=TRUE}
cbd.2 = ggplot(dat, aes(x = time / 60, y = beta, color = type, group = type)) +
  # geom_ribbon(aes(ymin = lower.joint, ymax = upper.joint, fill = type), 
  #             alpha = 0.3, color = NA) +  # Joint CI
  
  geom_ribbon(aes(ymin = lower.pointwise, ymax = upper.pointwise, fill = type),
              alpha = 0.3, color = NA) +  # Pointwise CI
  
  geom_hline(yintercept = 0, linetype = "dashed", color = "black") +
  
  geom_line(size = 1) +
  
  scale_color_brewer(palette = "Set1") +
  scale_fill_brewer(palette = "Set1") + 
  scale_x_continuous(
    breaks = c(1, 6, 12, 18, 23),
    labels = function(x) paste0(sprintf("%02d", x), ":00")
  ) +
  labs(title = 'Summarized Results (Include CI)',
       x = "Time (hour)", 
       y = 'Physical Intensity (count/min)',
       color = "Type", fill = "Type") +
  theme_minimal() +
  theme(
    panel.grid.major = element_line(color = "gray90", size = 0.2),
    axis.text = element_text(size = 12),
    axis.title = element_text(size = 12, face = "bold"),
    legend.title = element_text(size = 12, face = "bold"),
    legend.text = element_text(size = 10)
  )

print(cbd.2)
```

## Point Esitimate + 95% joint-wise CI

```{r}
cbd.3 = ggplot(dat, aes(x = time / 60, y = beta, color = type, group = type)) +
  geom_ribbon(aes(ymin = lower.joint, ymax = upper.joint, fill = type), 
              alpha = 0.3, color = NA) +  # Joint CI
  
  # geom_ribbon(aes(ymin = lower.pointwise, ymax = upper.pointwise, fill = type), 
  #             alpha = 0.3, color = NA) +  # Pointwise CI
  
  geom_hline(yintercept = 0, linetype = "dashed", color = "black") +
  
  geom_line(size = 1) +
  
  scale_color_brewer(palette = "Set1") +
  scale_fill_brewer(palette = "Set1") + 
  scale_x_continuous(
    breaks = c(1, 6, 12, 18, 23),
    labels = function(x) paste0(sprintf("%02d", x), ":00")
  ) +
  labs(title = 'Summarized Results (Include CI)',
       x = "Time (hour)", 
       y = 'Physical Intensity (count/min)',
       color = "Type", fill = "Type") +
  theme_minimal() +
  theme(
    panel.grid.major = element_line(color = "gray90", size = 0.2),
    axis.text = element_text(size = 12),
    axis.title = element_text(size = 12, face = "bold"),
    legend.title = element_text(size = 12, face = "bold"),
    legend.text = element_text(size = 10)
  )

print(cbd.3)
```

------------------------------------------------------------------------

## What We Plan to Do Next

1.  **Refinement of Functions** (December 2024 - January 2025):
    -   Improve current integration for single/multiple observations under various bootstrap methods.

    -   Incorporate the Unified Bootstrap method [@wang2022]

-   **Simulation Studies** (January - February 2025):

    -   Evaluate methods under different survey designs.

    -   Compare point-wise CI and JCB performance.

-   **Package Development & Manuscript Preparation** (February - March 2025):

    -   Package the functions for broader applications.

    -   Document findings and draft a manuscript.

-   **Thesis Submission** (April - May 2025):

    -   Finalize thesis incorporating NHATS analyses and simulation results.

------------------------------------------------------------------------

### NHATS Round 11 Study Design

The National Health and Aging Trends Study (NHATS) provides nationally representative data on physical activity patterns for Medicare beneficiaries aged 71 and older.

#### Accelerometry Subsample Design:

-   Initial Sample: 1,000 participants from Round 9 were flagged for accelerometry.

-   Eligible Sample:

    -   872 participants were eligible to wear the device.
    -   747 (86%) returned valid accelerometry data.

#### Data Collection:

-   Device: Actigraph CentrePoint Insight; Participants wore the device on their non-dominant wrist for 7 consecutive days (24 hours/day).
-   Processing: Raw accelerometry data were processed using the `ARCTOOLS` package.

#### Complex Survey Features:

-   Stratification (Stratum): `w11agvarstrat`.
-   Clustering (PSU): `w11agvarunit`
-   Full Sample Weights: `w11agfinwgt0` ensure unbiased population-level estimates.
    -   The full sample weight is for accelerometer data, which is adjusted for sampling probability and post-stratification

------------------------------------------------------------------------

## References:
